---
title: "Testing Mechanistic "
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf George~Kachergis (kachergis@stanford.edu)} \\ Department of Psychology, Stanford University \\ Stanford, CA
    \AND {\large \bf Martin~Zettersten (FIXME@princeton.edu)} \\ Department of Psychology, Princeton University \\ Princeton, NJ}

abstract: >
    Word learning is a complex process, influenced not only by the words and referents 
    in the language environment, but also by what learners selectively attend to,
    which in turn is likely determined by their prior knowledge. 
    
keywords: >
    active learning; cross-situational word learning; self-directed learning; uncertainty reduction
    
output: cogsci2016::cogsci_paper
#final-submission: \cogscifinalcopy
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
                      fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, 
                      message=F, sanitize = T)
```

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
```

```{r, load-data, echo=F}
load("data/cross_act_preprocessed.Rdata")

N_per_exp <- xd %>% 
  group_by(subject, experiment_name, age_group) %>%
  summarise(n = n()) %>%
  group_by(experiment_name, age_group) %>%
  summarise(n = n())
```


# Introduction

Early word learning is often thought of as largely determined by the learners'linguistic environment, where they can only learn what is present in the input.
Increasingly, research is finding that learners--even at a young age--are involved in selecting and shaping their input, and that such self-directed, active learning can accelerate acquisition. 
Understanding how learners choose their 

# Method

# Results


```{r 2-col-image, fig.env = "figure*", fig.pos = "h", fig.width=4, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "This image spans both columns. And the caption text is limited to 0.8 of the width of the document."}
img <- png::readPNG("figs/walrus.png")
grid::grid.raster(img)
```

## One-column images

Single column is the default option, but if you want set it explicitly, set `fig.env` to `figure`. Notice that the `num.cols` option for the caption width is set to `1`.

```{r image, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "One column image."}
img <- png::readPNG("figs/lab_logo_stanford.png")
grid::grid.raster(img)
```



## Tables

Number tables consecutively; place the table number and title (in
10 point) above the table with one line space above the caption and
one line space below it, as in Table 1. You may float
tables to the top or bottom of a column, set wide tables across both
columns.

You can use the xtable function in the xtable package.

```{r xtable, results="asis"}
n <- 100
x <- rnorm(n)
y <- 2*x + rnorm(n)
out <- lm(y ~ x)

tab1 <- xtable::xtable(summary(out)$coef, digits=c(0, 2, 2, 1, 2), 
                       caption = "This table prints across one column.")

print(tab1, type="latex", comment = F, table.placement = "H")
```

# Acknowledgements

Place acknowledgments (including funding information) in a section at
the end of the paper.

# References 

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent
