---
title: "Testing Mechanistic "
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf George~Kachergis (kachergis@stanford.edu)} \\ Department of Psychology, Stanford University \\ Stanford, CA
    \AND {\large \bf Martin~Zettersten (FIXME@princeton.edu)} \\ Department of Psychology, Princeton University \\ Princeton, NJ}

abstract: >
    Word learning is a complex process, influenced not only by the words and referents 
    in the language environment, but also by what learners selectively attend to,
    which in turn is likely determined by their prior knowledge. 
    
keywords: >
    active learning; cross-situational word learning; self-directed learning; uncertainty reduction
    
output: cogsci2016::cogsci_paper
#final-submission: \cogscifinalcopy
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
                      fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, 
                      message=F, sanitize = T)
```

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
require(tidyverse)
```

```{r, load-data, echo=F}
load("../data/cross_act_preprocessed.Rdata")

N_per_exp <- xd %>% 
  group_by(subject, experiment_name, age_group) %>%
  summarise(n = n()) %>%
  group_by(experiment_name, age_group) %>%
  summarise(n = n())
```


# Introduction

Early word learning is often thought of as largely determined by the learners' linguistic environment, where they can only learn what is present in the input.
Increasingly, research is finding that learners--even at a young age--are involved in selecting and shaping their input, and that such self-directed, active learning can accelerate acquisition. 
Understanding the strategies that learners use to shape their experience will help us build and improve theories of language.

The goal of our current study is to test how well previously proposed models of cross-situational word learning can explain the observed results of active learning experiments from @Zettersten2021.
Aside from assessing the ability of the models to account for the overall data, it is of particular interest whether particular model parameters are associated with individuals' 1) learning performance and 2) deployed active learning strategy.
We will also investigate age-related changes in model parameters.

# Method

## Data

All data were taken from @Zettersten2021. 
Data for our Study 1 were pooled from Experiment 1 (31 adults), Experiment 2 (40 children, 4-9 years of age), and adults in the Ambiguous Condition of Experiment S1 (FixMe adults), which had the same design as Experiments 1 and 2.

Data for Study 2 come from Experiment 3, comprising 58 children of ages 3 to 8.

## Models

We focused on analyzing data from the familiarity- and uncertainty-biased associative model [@Kachergis2012,@Kachergis2017], and variations of that model, described below.

### Familiarity- and Uncertainty-biased Associative Model

This model assumes that learners accumulate associations between all presented words and referents on each trial, but that their attention is biased more towards particular word-referent associations on the basis of prior experience.
In particular, more attention (i.e. associative weight) is directed to associations between words and referents that have appeared together on earlier trials (i.e., a familiarity bias).
In addition, more attention is directed to stimuli that have higher uncertainty (i.e. entropy, $H$) in their associations, whether because they are novel (and thus have no pre-existing associations) or because they have diffuse associations with several different stimuli.
The model assumes learners knowledge grows trial-by-trial, and that a fixed amount of associative weight (learning rate $\chi$) is distributed among the presented word-referent associations on a given trial in proportion to the prior familiarity of the association and the uncertainty of the involved word and referent.
The relative strength of the familiarity and uncertainty bias is governed by a parameter $\lambda$: for values of $\lambda>1$ the model focuses more attention on uncertain stimuli, while for values of $\lambda<1$ the model focuses more attention on familiar word-referent associations.
Associations are also assumed to decay from trial to trial, at a rate governed by parameter $\alpha$. 
The update rule for adjusting the association $M_{w,o}$ between word $w$ and referent $o$ on a given trial is

(update equation)

When tested on word $w$, the model assumes learners choose referent $o$ from the offered alternatives in proportion to their associative strength $M_{w,o}$ (i.e., Luce choice).

The other models we tested are variations on this model.
The strength-biased associative model implemented just the familiarity bias ...

### Uncertainty-biased Associative Model

### Novelty-biased Associative Model


## Fitting Procedure

We first fitted the model to each participant's data individually, 

# Results


```{r 2-col-image, fig.env = "figure*", fig.pos = "h", fig.width=4, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "This image spans both columns. And the caption text is limited to 0.8 of the width of the document."}
load("../data/ind_fits_kachergis_chi1.Rdata")
ind_fits <- ind_fits1

results <- ind_fits$fit_results %>% 
  mutate(age_years = ifelse(age<18, round(age), "adult"),
         age_years = ifelse(is.na(age_years), "adult", age_years)) %>% # 1 adult had NA age
  group_by(experiment_name, age_years, subject, age) %>%
  summarise(Human = mean(is_right),
            Model = mean(mod_perf)) %>%
  left_join(ind_fits$sfits) 

results %>% filter(is.element(experiment_name, c("Experiment 1","Experiment 2"))) %>%
  group_by(age_years) %>%
  summarise(hum_sd = sd(Human), 
            mod_sd = sd(Model),
            Human = mean(Human),
            Model = mean(Model)) %>%
  pivot_longer(cols = c(Human, Model), names_to = "group", values_to = "Accuracy") %>%
  ggplot(aes(x=age_years, y=Accuracy, color=group, shape=group)) + 
  geom_point() + # geom_linerange(aes(ymin = )) + 
  theme_classic() + xlab("Age (years)") + 
  ylim(0,1) + geom_hline(yintercept=1/8, linetype='dashed')
```

## One-column images

Single column is the default option, but if you want set it explicitly, set `fig.env` to `figure`. Notice that the `num.cols` option for the caption width is set to `1`.

```{r image, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "One column image."}

```



## Tables

Number tables consecutively; place the table number and title (in
10 point) above the table with one line space above the caption and
one line space below it, as in Table 1. You may float
tables to the top or bottom of a column, set wide tables across both
columns.

You can use the xtable function in the xtable package.

```{r xtable, results="asis"}
n <- 100
x <- rnorm(n)
y <- 2*x + rnorm(n)
out <- lm(y ~ x)

tab1 <- xtable::xtable(summary(out)$coef, digits=c(0, 2, 2, 1, 2), 
                       caption = "This table prints across one column.")

print(tab1, type="latex", comment = F, table.placement = "H")
```

# Acknowledgements

Place acknowledgments (including funding information) in a section at
the end of the paper.

# References 

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent
